---
title: "Homework 01"
author: "Nayla Trigueros Ortiz"
---

## TidyTuesday Section

Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge.
Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).
Provide sufficient background for readers to grasp your narrative.

### Research Question

This week's TidyTuesday brings insights into Brazil's companies, brought together by Marcelo Silva from Brazil's open CNPJ and the Brazilian Ministry of Finance.
For the assignment's purposes, I will focus on the following research question: What distinguishes Brazil's highest-capitalized companies from typical firms, and how do legal structure, ownership type, and formal size classifications shape access to significant capital?

### Data Import

```{r}
install.packages("tidytuesdayR")
tuesdata <- tidytuesdayR::tt_load('2026-01-27')
library(tidyverse)
library(dplyr)
library(ggplot2)

companies <- tuesdata$companies
legal_nature <- tuesdata$legal_nature
qualifications <- tuesdata$qualifications
size <- tuesdata$size
```

### Visualizations

```{r fig.width= 12, fig.height= 8}
# Data wrangling
companies_clean <- companies %>%
  filter(!is.na(capital_stock), !is.na(company_size), capital_stock > 0) %>%
  mutate(
    elite_threshold = quantile(capital_stock, 0.90),
    is_elite = capital_stock >= elite_threshold
  ) %>%
  
  # Get top 5 legal natures
  add_count(legal_nature, name = "legal_count") %>%
  mutate(legal_simple = if_else(
    legal_count >= sort(unique(legal_count), decreasing = TRUE)[5],
    legal_nature,
    "Other"
  ))

ggplot(companies_clean, aes(x = company_size, y = capital_stock, color = legal_simple)) +
  geom_jitter(alpha = 0.5, width = 0.2) +
  geom_hline(aes(yintercept = elite_threshold), linetype = "dashed", color = "red", size = 1) +
  scale_y_log10() +
  labs(
    title = "Capital Stock Across Company Sizes",
    subtitle = "Red line marks the top 10% threshold",
    x = "Company Size",
    y = "Capital Stock (BRL, log scale)",
    color = "Legal Nature"
  ) +
  scale_color_viridis_d() +
  theme_minimal()

# Data wrangling
owner_comparison <- companies %>%
  filter(!is.na(owner_qualification), !is.na(capital_stock), capital_stock > 0) %>%
  mutate(
    elite_threshold = quantile(capital_stock, 0.90),
    is_elite = if_else(capital_stock >= elite_threshold, "Elite (Top 10%)", "Other Companies")
  ) %>%
  
  # Get top 5 owner qualifications
  add_count(owner_qualification, name = "owner_count") %>%
  filter(owner_count >= sort(unique(owner_count), decreasing = TRUE)[5]) %>%
  group_by(owner_qualification, is_elite) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(is_elite) %>%
  mutate(percent = count / sum(count) * 100)

ggplot(owner_comparison, aes(x = reorder(owner_qualification, percent), y = percent, fill = is_elite)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    title = "Who Controls the Capital?",
    subtitle = "Owner qualification patterns in elite vs. typical Brazilian companies",
    x = "Owner Qualification",
    y = "Percentage (%)",
    fill = "Company Type"
  ) +
  scale_fill_viridis_d() +
  theme_minimal()
```

### Summary of Findings

Who controls the capital in Brazilian business?

Brazil's CNPJ registry tracks millions of companies, including their capital stock (declared share capital).
By examining the "capital elite" (top 10% by capital stock, above R\$3930456), we can understand who controls significant financial resources in Latin America's largest economy.

Does size matter for these businesses?

Our first visualization shows an interesting pattern: formal size classifications are nearly meaningless.
Companies labeled "micro-enterprises" can have capital stock exceeding R\$1 billion, often more than many "small-enterprises." The elite threshold (red line) cuts across all size categories, showing that Brazil's administrative labels don't reflect economic reality.

However, Legal structure does matter.
Limited Liability Business Companies dominate the high-capital tier, suggesting certain legal forms provide better access to serious investment.

What about ownership?

While Managing Partners control 79% of typical Brazilian firms, elite companies show much greater diversity in leadership.
Administrators and Managers lead 30.5% of elite companies but only 8.7% of typical firms.
Directors, Officers, and Presidents are also far more common at the top, suggesting that high-capital firms adopt more formalized corporate governance structures rather than the partnership model that dominates smaller enterprises.

So what does that mean?

Access to capital in Brazil isn't just about business ideas.
It's about structure.
Formal size classifications may misdirect policy support, while legal form and ownership type systematically advantage certain organizational arrangements.
For entrepreneurs seeking investment, these aren't just administrative choices.
They're strategic decisions with real consequences for accessing the capital that fuels growth.
In Brazil's business landscape, structure and ownership play a bigger role than size.
